---
title: "Torrential Twitter? Investigating the Severity of Harassment when Canadian Female Politicians Tweet about Climate Change"
author: "Inessa De Angelis"
thanks: "Code and other information available at: https://github.com/InessaDeAngelis/Torrential_Twitter_Trial "
date: "today"
date-format: "long"
abstract: "Women in politics face considerable online and offline harassment for their participation, especially when adopting strong policy stances on issues such as climate change. Known as gendertrolling, this specific form of harassment, in part, is due to male climate deniers feeling threatened by the perceived power and qualifications of female politicians. This paper analyzes Tweets which specifically discuss climate change from former MP Catherine McKenna, MP Elizabeth May, and MP Laurel Collins to investigate the severity of harassment on a seven-point scale and type of accounts behind the Tweets. This trial study reveals that nearly 80% of all Tweets contained some level of harassment, most frequently either questioning the authority of the MPs or calling them names or gender-based insults and came from accounts that were either suspended/deleted or personal users."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#### Workspace set-up ####

## Load packages ##
library(tidyverse)
library(knitr)
library(kableExtra)
library(here)
library(ggplot2)
library(formattable)
library(tayloRswift)

# Read in the cleaned McKenna data # 
finalized_mckenna_data <-
  read_csv(here::here(
      "~/Desktop/Final Paper/Outputs/Data/finalized_mckenna_data.csv"),
    show_col_types = FALSE
  )

# Read in the cleaned May data #
finalized_may_data <- 
  read_csv(here::here ("~/Desktop/Final Paper/Outputs/Data/finalized_may_data.csv"),
show_col_types = FALSE
)

# Read in the cleaned Collins data #
finalized_collins_data <- 
  read_csv(here::here ("~/Desktop/Final Paper/Outputs/Data/finalized_collins_data.csv"),
show_col_types = FALSE
)

# Create one combined data set with all three politicians #
combined_data <- rbind(finalized_mckenna_data, finalized_may_data, finalized_collins_data)
combined_data

# Create data set filtering bio data #
bio_data_MAGA <-
combined_data |>
  select(ID,
  name_of_politician,
  bio,
  severity_of_harassment, 
  type_of_account
  ) |>
  filter(grepl('MAGA', bio)) 

bio_data_PPC <-
combined_data |>
  select(ID,
  name_of_politician,
  bio,
  severity_of_harassment, 
  type_of_account
  ) |>
  filter(grepl('PPC', bio)) 

bio_data_CPC <-
combined_data |>
  select(ID,
  name_of_politician,
  bio,
  severity_of_harassment, 
  type_of_account
  ) |>
  filter(grepl('conservative', bio)) 

bio_data_CPC2 <-
combined_data |>
  select(ID,
  name_of_politician,
  bio,
  severity_of_harassment, 
  type_of_account
  ) |>
  filter(grepl('Conservative', bio)) 

bio_data_trudeau <-
combined_data |>
  select(ID,
  name_of_politician,
  bio,
  severity_of_harassment, 
  type_of_account
  ) |>
  filter(grepl('TrudeauMustGo', bio)) 

bio_data_ppc2 <-
combined_data |>
  select(ID,
  name_of_politician,
  bio,
  severity_of_harassment, 
  type_of_account
  ) |>
  filter(grepl('ppc', bio)) 

# Merge together all bio data sets #
combined_bio_data <- rbind(bio_data_CPC, bio_data_MAGA, bio_data_PPC, bio_data_trudeau, bio_data_CPC2, bio_data_ppc2)
combined_bio_data
  
```

# Introduction
Women in politics face considerable online and offline harassment, such as former Minister of the Environment and Climate Change Catherine McKenna and current Deputy Prime Minister Hon. Chrystia Freeland [@southern]. Known as “gendertrolling,” this harassment is part of a global trend of violence against women [@mantilla2013; @wagner]. Gender-based harassment on Twitter is well documented in academic literature and inexplicably pro-environmental behaviors are often viewed as “feminine” [@Citron; @mediatingmisogyny; @Anshelm]. Female politicians who adopt strong policy stances are more likely to face online misogyny and harassment because it is not consistent with existing gender stereotypes and notions of power, and poses a threat to male climate deniers [@courtemanche; @rheault; @degeus]

To gain a deeper understanding of the intersection of climate denialism and online gendered-harassment faced by Canadian female politicians, I measure the severity of harassment former Liberal Party MP Catherine McKenna, Green Party MP Elizabeth May, and New Democratic Party MP Laurel Collins receive when Tweeting about climate change on a seven-point scale, using Tweets sent leading up to and during the 2019 Federal election. I then analysed the types of accounts behind the Tweets sent to MP McKenna, MP May, and MP Collins. 

Harassment or gendertrolling is defined as sexist or misogynistic remarks that target a person based on their gender or sexuality and have varying levels of severity [@wagner]. Severity of harassment includes:

  1. Positive

  2. Neutral

  3. Questioning Authority 

  4. Name-calling/Gender insults

  5. Vicious language

  6. Credible threats

  7. Hate speech


Type of account is defined as the dynamics of relationships between followers, followings, and lists and will be determined by analyzing the accounts’ bio, profile image, and other characteristics [@singh; @uddin] and includes:

  1. Personal

  2. Professional

  3. Bots

  4. Spammers

  5. Anonymous 

  6. Suspended/Deleted

My analysis emphasizes that nearly 80% of the Tweets in my sample contained some level harassment, ranging from questioning authority to hate speech, while the remaining Tweets had either positive or neutral sentiments. Questioning the authority of female politicians or name-calling/offering up gender-based insults were the most common forms of harassment identified, accounting for 75% of all Tweets and aligning with expectations. Harassing Tweets were most likely to be sent from Suspended/deleted or personal Twitter accounts. My [Discussion] section further explore these key findings and others of note. 

In the remainder of this paper, I commence with the [Data] section which outlines the nature of the data used for this trial study, limitations, coding, cleaning, and modelling procedures. The [Model] section outlines and justifies the model used in this paper. In the [Results] section, I present trends found during the analysis process and results from my model. Then, in the [Discussion] section, I provide further insights and outline areas where my thesis will differ from this trial study. Finally, in the [Conclusion], I summarize the main findings. 

## Theoretical Framework
This study will be conducted under the theory of deliberative democracy which emphasizes having open, respectful, and diverse discourse for a healthy democracy [@pmc2019; @trifiro]. Democratic legitimacy is established by allowing the public to collectively participate in deliberation about consequential decisions, while recognizing there are multiple viewpoints on every issue, including climate change [@mckaytenove; @masullochen]. 

Results from the analyzed Tweets from the three female politicians will be compared to the aforementioned definition to see if the level of discourse about climate change on Twitter matches the characteristics of what is considered healthy discourse under the theory of deliberative democracy. 

# Data 
The data used in this paper was collected via Twitter's API in early 2021 for a separate research project I undertook during my undergraduate degree and has since been stored on my personal Mac. A more detailed discussion of the dataset composition and the collection process can be found in the [Appendix A]. 

## Data Limitations
Twitter, the platform's affordances, and the types of accounts and content circulate is ever-changing. Consequently, Tweets authored in 2019, collected in 2021, and analyzed for this trial study in 2023 are faced with changing contexts and affordances which must be accounted for. There are potential inaccuracies in reporting which accounts are still on the platform and which have been suspended/deleted, especially following successive waves of Covid-19 mis- and dis-information on the platform. Although this was anticipated, some users do not have bios or locations indicated (listed as NA), which may impact how `type of account` is coded. The sheer volume of Tweets collected in 2021 also means that only a small sample is being used for this trial study, which means there is a higher chance of inaccuracies and ability to not establish patterns to see the larger picture. 

During the 2019 Federal election campaign, MP Collins was a City Councillor in Victoria, BC, never having been elected to the House of Commons or faced the scrutiny of traditional and social media on a national level. This means that she received far fewer Tweets during the same period as MP McKenna and MP May and that the results from her study are far less generalizable, but are still important to include given her increased profile in 2023 and anticipated prominence in the full-scale version of this study. 

A major limitation of using Twitter data for any research project, including this one, is the inability to determine a representative sample of voting-age Canadians and draw generalizable results about public opinion on any given issue [@bermingham]. Selection bias exists because Twitter tends to only represent the views of people who are “... more partisan, polarized, and uncivil” and motivated to be active users [@mcgregor, p. 237]. We are also limited by the information users choose to publicly display on their profiles, such as their location, so it is not always possible to confirm if an individual is Canadian or other identifying demographic information [@holmberg]. 

Moreover, Twitter data is collected as a snapshot in time, often reducing users to a single interaction on the platform and leading to potential measurement bias. Consequently, with context collapse which happens when an individual user’s Tweets are taken out of the context and imagined audience in which they were intended for, it can be easy to misrepresent their true political beliefs, views on climate change, and perspectives on the qualifications of women in politics [@contextcollapse]. Some users may also show up more than once in the dataset if they are prolific Twitter users, leading to overrepresentation of their views and type of account. With these identified limitations in mind, generalizations from this trial study can only be reflective of the users in the dataset and not the wider Canadian public. 


## Data Coding
This trial study employs a novel approach of conducting quantitative content analysis, combining the level of nuance and detail captured in qualitative content analysis and interviews [@wagner; @erikson] with more binary quantitative methods [@theocharis; @rheault]. By offering this quantitative, seven-point scale to measure the severity of harassment, we can begin to better understand the full extent of harassment faced by Canadian female politicians and implement solutions. 

To code the Tweets, I designed a Google Form (available in [Appendix B]) which allows coders to enter key information about each Tweet and the type of account behind the Tweet. Coders code `severity of harassment` on the previously discussed seven-point scale based on the Tweet text provided in the dataset. `Type of account` was determined by coders by looking at a user's username, display name, and bio displayed in the dataset and then by looking at a user's profile and header pictures, frequency of replies and likes, as it stands of July 2023 on Twitter. Although there is a significant amount of time between when the Tweet was originally authored, collected, and coded for this project, it was determined that this was the best approach for this trial study, but will be re-evaluated for the full-scale study.   

## Data Cleaning
Data was collected via Twitter API using Python (see datasheet in [Appendix A] for more information) and cleaned, tested, modeled, and analyzed using the statistical programming software R [@r], using functions from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `tayloRswift` [@taylor], `janitor` [@rJanitor], `KableExtra` [@rKableExtra], `knitr` [@rknitr], `here` [@rHere], `formattable` [@rFormattable], `arrow` [@arrow], `validate` [@validate], `testthat` [@testthat], `rstanarm` [@rstanarm], `marginaleffects` [@marginal], and `modelsummary` [@modelsummary].

The data utilized in this trial study underwent two rounds of cleaning: the first was to clean and select the Tweets to manually code and the second was to clean the coded Tweets to prepare for analysis and visualization. 

### Raw Data 
In preparation to code my Tweets for severity of harassment and type of account, I first had to eliminate extra information in the data sets and take a sample, given that there were 116,649 Tweets between the three politicians in the raw datasets. 

After retrieving the data sets from my archived files, I performed the first round of data cleaning based on the design of my trial study. I removed the columns which presented information not required for this study and using the `filter` function, I filtered all Tweets by the key word "climate", and then took a sample using the `slice_sample` function of 500 Tweets from both MP McKenna and MP May's data sets. Only 13 Tweets from MP Collins matched the criteria, so there was no need to take a sample. 

I renamed columns and capitalized the first letter of every word using Twitter Developer's guidelines and my Codebook [@Twittermedia]. I renamed `text` to `Text`, `user_name` to `Name`, `user_sname` to `Username`, `user_description` to `Bio`, `user_location` to `Location`, `followers` to `Followers`, `friends` to `Following`, and `tweet_URL` to `URL`. 

I then tested the raw data using the `testthat` [@testthat] package to ensure the data set was ready to be manually coded.

### Coded Data 
After coding the 1,013 Tweets manually using my Google form (see [Appendix B]), I then visually checked for missing data, such as Tweet URLs, severity of harassment, and type of account and went back and re-coded the few Tweets that contained missing information. A number of Twitter accounts opted not have bios (listed as NA in the data sets), which is missing data that was anticipated within the data sets. After my data set passed this initial, informal test, I proceeded to clean it a second time using R. 

For my `severity of harassment` data, I updated value 1 to "Positive", 2 to "Neutral", 3 to "Questioning authority", 4 to "Name-calling/Gender insults", 5 to "Vicious language", 6 to "Credible threats", and 7 to "Hate speech" based on my Codebook. 

For my `type of account` data, I changed value 1 to "Personal", 2 to "Professional", 3 to "Bots", 4 to "Spammers", 5 to "Anonymous", and 6 to "Suspended/deleted" based on my Codebook. 

I tested the cleaned data using the `validate` [@validate] package, primarily validating variable types and ensuring all Tweets were coded (500 for both MP McKenna and MP May and 13 for MP Collins). 

# Model

## Model Justification 

# Results
## Severity of Harassment
Among the 1,013 Tweets selected for content analysis from the three MPs, 522 (51.5%) Tweets were identified as **Questioning authority**, 243 (24.0%) as **Name-calling/Gender insults**, and 208 (20.5%) as **Neutral**. 24 (2.4%) Tweets were coded as **Positive**, 12 (1.2%) as **Vicious language**, while 2 Tweets (or 0.2%) were identified for both the **Credible threats** and **Hate speech** categories (see @tbl-overall_severity). 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-overall_severity
#| tbl-cap: Breakdown of Severity of Harassment on the seven-point scale detected in Tweets to all three politicians

# Code to arrange severity of harassment in order from 1-7 referenced from: https://www.statology.org/dplyr-arrange-custom-order/

combined_data |>
  group_by(severity_of_harassment) |>
  count() |>
  ungroup() |>
  mutate (pct = percent(n / sum(n), digit = 1)) |>
  arrange(match(severity_of_harassment, c('Positive', 'Neutral', 'Questioning Authority', 'Name-calling/Gender insults', 'Vicious language', 'Credible threats', 'Hate speech'))) |>
  kable(
    col.names = c("Severity of Harassment", "Number of Responses", "Percent"),
    booktabs = TRUE,
  )

```

@fig-severity_all illustrates that the `severity of harassment` detected across the Tweets mentioning all three female politicians have some commonalities, but also differences depending on the politician. Just over half of both MP McKenna and MP May's Tweets were categorized as **Questioning Authority**, while MP Collins only had 2 Tweets that questioned her authority. The next most frequent level of harassment detected for MP McKenna was **Name-calling/Gender insults** which accounted for 30% of all Tweets, while MP May's next most frequent type of harassment was **Neutral** at almost 30%. In contrast, the third most frequent level of harassment detected for MP McKenna was **Neutral**, with 10% of Tweets, while MP May's next most frequent level of harassment was **Name-calling/Gender insults**, accounting for almost 20% of Tweets. MP Collins' remaining 85% of Tweets were categorized as **Neutral**. 

For both MP McKenna and MP May, the remaining levels of harassment including **Positive**, **Vicious language**, **Credible threats**, and **Hate Speech** differed to some degree, as emphasized by @fig-severity_all. 2% of MP McKenna's Tweets were **Positive** while 3% of the Tweets sent to MP May had a **Positive** tone. 2% of MP McKenna's Tweets contained **Vicious language** while less than 1% of the Tweets sent to MP May contained **Vicious language**. **Credible threats** and **Hate Speech** accounted for less than 1% of the Tweets sent to MP McKenna, with MP May having no **Credible threats** detected among the Tweets sent to her and less than 1% containing **Hate Speech**. 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-severity_all
#| fig-cap: Breakdown of Severity of Harassment on the seven-point scale detected in Tweets by politician

# Code to order severity of harassment according to seven-point scale referenced from: https://r-graph-gallery.com/267-reorder-a-variable-in-ggplot2.html

combined_data |>
  mutate(severity_of_harassment = fct_relevel(severity_of_harassment, "Positive", "Neutral", "Questioning Authority", "Name-calling/Gender insults", "Vicious language", "Credible threats", "Hate speech")) |>
  ggplot(mapping = aes(x = severity_of_harassment, fill = name_of_politician)) +
  geom_bar(position = "dodge2") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle=45, hjust = 1, size = 10)) + 
  labs(x = "Severity of Harassment", y = "Number of Tweets", fill = "Politician") +
  theme(legend.position = "bottom") +
  theme(legend.text = element_text(size = 6)) +
  theme(legend.title = element_text(size = 9)) +
  scale_fill_taylor(palette = "lover",
  discrete = TRUE,
  reverse = FALSE)
   
```

## Type of Account
Of the accounts behind the Tweets sent to the three politicians, 422 (41.7%) were identified as **Suspended/deleted**, 327 (32.3%) as **Personal**, and 150 (14.8%) as **Anonymous**. 102 (10.1%) were coded as **Professional**, 11 (1.1%) as **Spammers**, while 1 account (or 0.1%) was identified as a **Bot** (see @tbl-overall_account). 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-overall_account
#| tbl-cap: Breakdown of Types of Accounts detected in Tweets from all three politicians

combined_data |>
  group_by(type_of_account) |>
  count() |>
  ungroup() |>
  mutate (pct = percent(n / sum(n), digit = 1)) |>
  arrange(match(type_of_account, c('Personal', 'Professional', 'Bots', 'Spammers', 'Anonymous', 'Suspended/deleted'))) |>
  kable(
    col.names = c("Type of Account", "Number of Responses", "Percent"),
    booktabs = TRUE,
  )

```

@fig-accounts_all illustrates that the `types of accounts` detected as being behind the Tweets sent to the three female politicians share some common features, but also differences depending on the politician. Just over 40% of Tweets sent to MP McKenna came from **Suspended/deleted** users while just under 40% of Tweets sent to MP May came from **Suspended/deleted** users. Approximately 25% of Tweets sent to MP McKenna came from a **Personal** account while nearly 40% of Tweets sent to MP May came from a **Personal** account. **Anonymous** accounts were the third most frequent type of account detected for MP McKenna, sending just over 20% of the Tweets in her sample. Conversely, the third most frequent type of account detected for MP May was **Professional**, responsible for sending just under 15% of the Tweets in her sample. The `types of accounts` responsible for sending Tweets to MP Collins were **Professional** users (70%), followed by **Personal** users (almost 25%), and then **Suspended/deleted** users (just under 5%). 

For the fourth most common `type of accounts` identified, just over 5% of Tweets sent to MP McKenna came from **Professional** accounts while MP May received just under 10% of Tweets from **Anonymous** accounts (see @fig-accounts_all). The remaining 2% of Tweets received by MP McKenna came from **Spammers**, while **Spammers** and **Bots** combined were responsible for sending less than 1% of the remaining Tweets to MP May. Both MP McKenna and Collins had no **Bots** detected. 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-accounts_all
#| fig-cap: Type of Accounts Behind the Tweets Received by All Three Politicians

combined_data |>
  mutate(type_of_account = fct_relevel(type_of_account, "Personal", "Professional", "Bots", "Spammers", "Anonymous", "Suspended/deleted")) |>
  ggplot(mapping = aes(x = type_of_account, fill = name_of_politician)) +
  geom_bar(position = "dodge2") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle=45, hjust = 1, size = 10)) + 
  labs(x = "Type of Account", y = "Number of Tweets", fill = "Politician") +
  theme(legend.position = "bottom") +
  theme(legend.text = element_text(size = 6)) +
  theme(legend.title = element_text(size = 9)) +
  scale_fill_taylor(palette = "lover",
  discrete = TRUE,
  reverse = FALSE)
   
```

## Catherine McKenna
Among the sample of 500 Tweets sent to MP McKenna, 270 (54%) were identified as **Questioning authority**, 153 (30.6%) as **Name-calling/Gender insults**, and 54 (10.8%) as **Neutral** for `severity of harassment`. 10 Tweets (or 2.0%) were identified for both the **Positive** and **Vicious language** categories, while 2 Tweets (0.4%) contained **Credible threats** and 1 Tweet (0.2%) contained **Hate speech** (see @tbl-mckenna_severity). 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-mckenna_severity
#| tbl-cap: Breakdown of the Severity of Harassment on the seven-point scale detected among MP McKenna's Tweets

finalized_mckenna_data |>
  group_by(severity_of_harassment) |>
count() |>
  ungroup() |>
  mutate(pct = percent(n / sum(n), digit = 1)) |>
  arrange(match(severity_of_harassment, c('Positive', 'Neutral', 'Questioning Authority', 'Name-calling/Gender insults', 'Vicious language', 'Credible threats', 'Hate speech'))) |>
  kable(
    col.names = c("Severity of Harassment", "Number of Responses", "Percentage"),
    booktabs = TRUE,
  )

```

For `Type of account`, 222 (44.4%) were detected as **Suspended/deleted**, 131 (26.2%) identified as **Personal**, 109 (21.8%) as **Anonymous**, 28 (5.6%) as **Professional**, 10 (2.0%) as **Spammers**, with no **bots** being detected (see @tbl-mckenna_account). 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-mckenna_account
#| tbl-cap: Breakdown of the Types of Accounts detected among MP McKenna's Tweets

finalized_mckenna_data |>
  group_by(type_of_account) |>
  count() |>
  ungroup() |>
  mutate (pct = percent(n / sum(n), digit = 1)) |>
  arrange(match(type_of_account, c('Personal', 'Professional', 'Bots', 'Spammers', 'Anonymous', 'Suspended/deleted'))) |>
  kable(
    col.names = c("Type of Account", "Number of Responses", "Percentage"),
    booktabs = TRUE,
  )

```

@fig-mckenna_relationship emphasizes the relationship between the `severity of harassment` faced by MP McKenna and the `type of accounts` behind the Tweets. The few **Positive** Tweets received by MP McKenna were most likely to be sent by **Personal** users (1%), **Professional** users (0.6%), or **Suspended/deleted** users (0.4%), while **Neutral** Tweets came from accounts categorized as **Professional** (3.8%), **Suspended/deleted** (3.6%), and **Personal** (2.4%). **Anonymous** users and **Spammers** did not send **Positive** Tweets to MP McKenna. Tweets coded as **Questioning Authority** were most frequently sent by **Suspended/deleted** users (22.6%), followed by **Personal** users (15.6%) and **Anonymous** users (13.4%). The Tweets containing **Name-calling/Gender insults** were most frequently sent by users identified as **Suspended/deleted** (16.2%), **Anonymous** (7.0%), or **Personal** (6.8%). **Vicious language** identified Tweets were sent by **Suspended/deleted** users (1.2%), **Anonymous** users (0.4%), and **Spammers** (0.2%). The identified instances of **Credible threats** and **Hate speech** were sent by **Suspended/deleted** accounts and **Bots** did not send any Tweets to MP McKenna. 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-mckenna_relationship
#| fig-cap: Relationship between Severity of Harassment and Types of Accounts detected among MP McKenna's Tweets

finalized_mckenna_data |>
  mutate(severity_of_harassment = fct_relevel(severity_of_harassment, "Positive", "Neutral", "Questioning Authority", "Name-calling/Gender insults", "Vicious language", "Credible threats", "Hate speech")) |>
  mutate(type_of_account = fct_relevel(type_of_account, "Personal", "Professional", "Bots", "Spammers", "Anonymous", "Suspended/deleted")) |>
  ggplot(mapping = aes(x = severity_of_harassment, fill = type_of_account)) +
  geom_bar() +
  theme_minimal() +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 13)) +
  theme(axis.text.x = element_text(angle=45, hjust = 1, size = 9)) + 
  theme(axis.text.y = element_text(size = 6)) +
  labs(x = "Severity of Harassment", y = "Number of Tweets", fill = "Type of Account") +
  theme(legend.position = "bottom") +
  theme(legend.text = element_text(size = 6)) +
  theme(legend.title = element_text(size = 9)) +
  scale_fill_taylor(palette = "lover",
  discrete = TRUE,
  reverse = TRUE)
 
```

## Elizabeth May
Among MP May's sample of 500 Tweets, 250 (50.0%) were identified as **Questioning authority**, 143 (28.6%) as **Neutral**, and 90 (18.0%) as **Name-calling/Gender insults** for `severity of harassment`. 14 Tweets (2.8%) were identified as being **Positive**, while 2 Tweets (0.4%) contained **Vicious language** and 1 Tweet (0.2%) contained **Hate speech** (see @tbl-may_severity). No Tweets containing **Credible threats** were detected in this sample.

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-may_severity
#| tbl-cap: Breakdown of the Severity of Harassment on the seven-point scale detected among MP May's Tweets

finalized_may_data |>
  group_by(severity_of_harassment) |>
  count() |>
  ungroup() |>
  mutate (pct = percent(n / sum(n), digit = 1)) |>
  arrange(match(severity_of_harassment, c('Positive', 'Neutral', 'Questioning Authority', 'Name-calling/Gender insults', 'Vicious language', 'Credible threats', 'Hate speech'))) |>
  kable(
    col.names = c("Severity of Harassment", "Number of Responses", "Percentage"),
    booktabs = TRUE,
  )

```

For `Type of account`, 199 (39.8%) were detected as **Suspended/deleted**, 193 (38.6%) identified as **Personal**, 65 (13.0%) as **Professional**, and 41 (8.2%) as **Anonymous**. 1 (0.2%) of each **Bots** and **Spammers** were detected (see @tbl-may_account). 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-may_account
#| tbl-cap: Breakdown of the Type of Accounts detected among MP May's Tweets

finalized_may_data |>
  group_by(type_of_account) |>
  count() |>
  ungroup() |>
  mutate (pct = percent(n / sum(n), digit = 1)) |>
  arrange(match(type_of_account, c('Personal', 'Professional', 'Bots', 'Spammers', 'Anonymous', 'Suspended/deleted'))) |>
  kable(
    col.names = c("Type of Account", "Number of Responses", "Percentage"),
    booktabs = TRUE,
  )

```

The relationship between the `severity of harassment` faced by MP May and the `type of accounts` behind the Tweets can be observed in @fig-may_relationship. **Positive** Tweets were sent by either **Personal** users (1.4%) or **Professional** users (0.4%), while **Neutral** Tweets most frequently came from accounts categorized as **Personal** (10.9%), **Professional** (9.4%), and **Suspended/deleted** (7.4%). Tweets coded as **Questioning Authority** were most frequently sent by **Suspended/deleted** users (22.2%), followed by **Personal** users (20.2%) and **Anonymous** users (4.2%). Tweets **Questioning Authority** was the only instance in which **Bots** and **Spammers** were detected in MP May's sample. The Tweets containing **Name-calling/Gender insults** were most frequently sent by users identified as **Suspended/deleted** (8.8%), **Personal** (6.4%), or **Anonymous** (2.6%). Moreover, **Vicious language** identified Tweets were sent by **Suspended/deleted** users (0.2%) and **Anonymous** users (0.2%). There were no identified instances of **Credible threats** and the 1 instance of **Hate speech** was sent by a **Suspended/deleted** account. 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-may_relationship
#| fig-cap: Relationship between Severity of Harassment and Types of Accounts detected among MP May's Tweets

finalized_may_data |>
  mutate(severity_of_harassment = fct_relevel(severity_of_harassment, "Positive", "Neutral", "Questioning Authority", "Name-calling/Gender insults", "Vicious language", "Credible threats", "Hate speech")) |>
  mutate(type_of_account = fct_relevel(type_of_account, "Personal", "Professional", "Bots", "Spammers", "Anonymous", "Suspended/deleted")) |>
  ggplot(mapping = aes(x = severity_of_harassment, fill = type_of_account)) +
  geom_bar() +
  theme_minimal() +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 13)) +
  theme(axis.text.x = element_text(angle=45, hjust = 1, size = 9)) + 
  theme(axis.text.y = element_text(size = 6)) +
  labs(x = "Severity of Harassment", y = "Number of Tweets", fill = "Type of Account") +
  theme(legend.position = "bottom") +
  theme(legend.text = element_text(size = 6)) +
  theme(legend.title = element_text(size = 9)) +
  scale_fill_taylor(palette = "lover",
  discrete = TRUE,
  reverse = TRUE)

```

## Laurel Collins
Among this small sample of 13 Tweets sent to MP Collins, 11 (84.6%) were identified as **Neutral** and 2 (15.4%) as **Questioning authority** for `severity of harassment` (see @tbl-collins_severity). No Tweets containing **Positive**, **Name-calling/Gender insults**, **Vicious language**, **Credible threats**, or **Hate speech** were detected in this sample.

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-collins_severity
#| tbl-cap: Breakdown of the Severity of Harassment on the seven-point scale detected among MP Collins' Tweets

finalized_collins_data |>
  group_by(severity_of_harassment) |>
  count() |>
  ungroup() |>
  mutate (pct = percent(n / sum(n), digit = 1)) |>
  arrange(match(severity_of_harassment, c('Positive', 'Neutral', 'Questioning Authority', 'Name-calling/Gender insults', 'Vicious language', 'Credible threats', 'Hate speech'))) |>
  kable(
    col.names = c("Severity of Harassment", "Number of Responses", "Percentage"),
    booktabs = TRUE,
  )

```

For `Type of account`, 9 (69.2%) were detected as **Professional**, 3 (23.1%) as **Personal**, and 1 (7.7%) as **Suspended/deleted** (see @tbl-collins_account). No **Bots**, **Spammers**, or **Anonymous** accounts were detected in this sample. 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-collins_account
#| tbl-cap: Breakdown of the Types of Accounts detected among MP Collins' Tweets

finalized_collins_data |>
  group_by(type_of_account) |>
  count() |>
  ungroup() |>
  mutate (pct = percent(n / sum(n), digit = 1)) |>
  arrange(match(type_of_account, c('Personal', 'Professional', 'Bots', 'Spammers', 'Anonymous', 'Suspended/deleted'))) |>
  kable(
    col.names = c("Type of Account", "Number of Responses", "Percentage"),
    booktabs = TRUE,
  )

```

The relationship between `severity of harassment` and `types of accounts` present in MP Collins' sample of Tweets is available in @fig-collins_relationship. **Neutral** Tweets were sent by **Professional** users (61.5%), **Personal** users (15.4%), or **Suspended/deleted** users (7.7%). Tweets **Questioning Authority** were identified as being sent by either **Personal** users (7.7%) or **Professional** users (7.7%). No **Positive**, **Name-calling/Gender insults**, **Vicious language**, **Credible threats**, or **Hate speech** Tweets were detected in this sample and neither were **Bots**, **Spammers**, or **Anonymous** accounts. 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-collins_relationship
#| fig-cap: Relationship between Severity of Harassment and Types of Accounts detected among MP Collin's Tweets

finalized_collins_data |>
  mutate(severity_of_harassment = fct_relevel(severity_of_harassment, "Positive", "Neutral", "Questioning Authority", "Name-calling/Gender insults", "Vicious language", "Credible threats", "Hate speech")) |>
  mutate(type_of_account = fct_relevel(type_of_account, "Personal", "Professional", "Bots", "Spammers", "Anonymous", "Suspended/deleted")) |>
  ggplot(mapping = aes(x = severity_of_harassment, fill = type_of_account)) +
  geom_bar() +
  theme_minimal() +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme(axis.text.x = element_text(angle=45, hjust = 1, size = 9)) + 
  theme(axis.text.y = element_text(size = 6)) +
  labs(x = "Severity of Harassment", y = "Number of Tweets", fill = "Type of Account") +
  theme(legend.position = "bottom") +
  theme(legend.text = element_text(size = 6)) +
  theme(legend.title = element_text(size = 9)) +
  scale_fill_taylor(palette = "lover",
  discrete = TRUE,
  reverse = FALSE)
  
```
## Model 

# Discussion
Overall, the data from this trial study reveals some form of harassment is present in over two-thirds of the Tweets sent to MP McKenna, MP May, and MP Collins which explicitly discuss climate change (see @tbl-overall_severity). The `severity of harassment` and `type of accounts` behind the Tweets vary by female politician, with MP McKenna seeing the more severe forms of harassment and MP Collins seeing the least. The following sections will further discuss these identified trends and compared them to the characteristics of what constitutes a healthy deliberative democracy. 

## Severity of Harassment
As hypothesized, the `severity of harassment` detected in Tweets for this trial study varies on the seven-point scale, with at least one Tweet being detected for each level (see @tbl-overall_severity).

I also hypothesized that harassing Tweets were more likely to be categorized as **Questioning Authority**, **Name-calling/gender insults**, and/or **Vicious language**. My hypothesis was two-thirds correct, with 51.5% of Tweets coded as **Questioning Authority** and 24.0% coded as **Name-calling/gender insults**. However, only 1.2% of Tweets were categorized as **Vicious language**, with **Neutral** (at 20.5%) being the third most identified level in this trial study (further details available in @tbl-overall_severity). Although this hypothesis was not entirely validated through this trial study, I anticipate it being proven in the full-scale version of this study for reasons explained below. 

This discrepancy between **Vicious language** and **Neutral** can be explain by a few factors. Firstly, stronger content moderation processes were applied to content on Twitter prior to Elon Musk's purchase of the platform in 2022, leading to Tweets with more profanities being removed at a higher rate, which will likely change with Musk's promotion of "free speech" in 2023 [@Citron]. Secondly, I did not anticipate there being as much civil discourse about climate change and environmental policy as there was on/around the accounts of MP May and MP Collins. It is difficult to be certain whether this will be the case in the full-scale version of this study, with the more polarized nature of Canadian politics following the Covid-19 pandemic and MP Collins' raised profile which is often a cause of more severe forms of discourse and harassment [@rheault]. 

## Type of Account
As hypothesized, at least one account was identified for each of the six `types of accounts`. **Suspended/deleted** accounts constituted 41.7% of all accounts, followed by **Personal** (32.3%) and **Anonymous** (14.8%) (a full breakdown is available in @tbl-overall_account). The number of **Personal** accounts approximately aligns with number I anticipated finding in this trial study. However, I did not anticipate as many as 41.7% of Tweets would be authored by **Suspended/deleted** users, which could in part be due to the length of time which has passed since these Tweets were authored and collected (further discussed in [Limitations]). 

Moreover, I suspected that the more severe forms of harassment identified in this trial study would come from **Personal** and **Suspended/deleted accounts**. Given the low number of Tweets containing **Vicious language**, **Credible threats**, and **Hate speech**, this hypothesis will be applied to the results for **Questioning Authority** and **Name-calling/Gender insults**. **Personal** users accounted for 17.8% of the Tweets **Questioning Authority**, while **Suspended/deleted accounts** sent 22.1% of the Tweets. **Personal** accounts were responsible for sending 6.5% of the Tweets containing **Name-calling/Gender insults**, while **Suspended/deleted accounts** sent 12.3% (see @tbl-type_hypothesis). 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-type_hypothesis
#| tbl-cap: Breakdown of the Types of Accounts detected among MP Collins' Tweets

combined_data |>
  group_by(severity_of_harassment, type_of_account) |>
  count() |>
  ungroup() |>
  mutate (pct = percent(n / sum(n), digit = 1)) |>
  arrange(match(severity_of_harassment, c('Questioning Authority', 'Name-calling/Gender insults'))) |>
  filter(severity_of_harassment == 'Questioning Authority' | severity_of_harassment == 'Name-calling/Gender insults') |>
  filter(type_of_account == 'Personal' | type_of_account == 'Suspended/deleted') |>
  kable(
    col.names = c("Severity of Harassment", "Type of Account", "Number of Responses", "Percent"),
    booktabs = TRUE,
  )

```

## Severity of Harassment & Political Identification 
As hypothesized, accounts belonging to right-wing, ideologically driven users will sent Tweets containing harassing content to the three female politicians. This data was gleaned by employing key word selection functions for users' bios. 114 accounts had information in their bios that emphasized their right-wing political views. 

Unsurprisingly, @tbl-con_severity illustrates that 64 (56.1%) of the Tweets were **Questioning authority**, 45 (39.5%) were **Name-calling/Gender insults**, 4 (3.5%) were **Neutral**, and 1 (0.9%) contained **Vicious language**. No Tweets were coded as **Positive**, **Credible threats**, or **Hate speech** which emphasizes that users with right-wing views consistently felt motivated enough to engage with MP McKenna and MP May on Twitter, but not on either extreme of the seven-point scale. 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-con_severity
#| tbl-cap: Breakdown of the Type of Accounts detected among accounts that hold right-wing views and Severity of Harassment 

combined_bio_data |>
  group_by(severity_of_harassment) |>
  count() |>
  ungroup() |>
  mutate (pct = percent(n / sum(n), digit = 1)) |>
  arrange(match(severity_of_harassment, c('Positive', 'Neutral', 'Questioning Authority', 'Name-calling/Gender insults', 'Vicious language', 'Credible threats', 'Hate speech'))) |>
  kable(
    col.names = c("Severity of Harassment", "Number of Tweets", "Percentage"),
    booktabs = TRUE,
  )

```

We can see in @tbl-con_accounts the breakdown of the different types of accounts that belong to people with right-wing views. 65 (57.0%) of the accounts are **Suspended/deleted**, while 32 (28.1%) are **Personal** and 13 (11.4%) are **Anonymous**. 2 (1.8%) of the accounts **Spammers** and similarly, 2 (1.8%) of the accounts are **Professional**. No **Bots** were identified. 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-con_accounts
#| tbl-cap: Breakdown of the Type of Accounts detected among accounts that hold right-wing views

combined_bio_data |>
  group_by(type_of_account) |>
  count() |>
  ungroup() |>
  mutate (pct = percent(n / sum(n), digit = 1)) |>
  arrange(match(type_of_account, c('Personal', 'Professional', 'Bots', 'Spammers', 'Anonymous', 'Suspended/deleted'))) |>
  kable(
    col.names = c("Type of Account", "Number of Accounts", "Percentage"),
    booktabs = TRUE,
  )

```


```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-accounts_ideology 
#| fig-cap: Relationship between Severity of Harassment, Type of Account, and Political Identification by politician

combined_bio_data |>
  mutate(severity_of_harassment = fct_relevel(severity_of_harassment, "Positive", "Neutral", "Questioning Authority", "Name-calling/Gender insults", "Vicious language", "Credible threats", "Hate speech")) |>
  mutate(type_of_account = fct_relevel(type_of_account, "Personal", "Professional", "Bots", "Spammers", "Anonymous", "Suspended/deleted")) |>
  ggplot(mapping = aes(x = name_of_politician, fill = type_of_account)) +
  geom_bar(position = 'dodge2') +
  theme_minimal() +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 3)) +
  theme(axis.text.x = element_text(angle=45, hjust = 1, size = 6)) + 
  theme(axis.text.y = element_text(size = 6)) +
  labs(
    x = "Politician",
    y = "Severity of Harassment",
    fill = "Type of Account"
  ) +
  theme(legend.position = "bottom") +
  theme(legend.text = element_text(size = 6)) +
  theme(legend.title = element_text(size = 9)) +
  scale_fill_taylor(palette = "lover",
  discrete = TRUE,
  reverse = TRUE) +
  facet_wrap(
    vars(severity_of_harassment),
    scales = "free",
  )
   
```

As shown in @fig-accounts_ideology_all, questioning authority and name-calling/gender insults were most frequently Tweeted by people with suspsended/deleted accounts, followed by personal accounts. 

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-accounts_ideology_all 
#| fig-cap: Relationship between Severity of Harassment, Type of Account, and Political Ideology 

combined_bio_data |>
  mutate(severity_of_harassment = fct_relevel(severity_of_harassment, "Positive", "Neutral", "Questioning Authority", "Name-calling/Gender insults", "Vicious language", "Credible threats", "Hate speech")) |>
  mutate(type_of_account = fct_relevel(type_of_account, "Personal", "Professional", "Bots", "Spammers", "Anonymous", "Suspended/deleted")) |>
  ggplot(mapping = aes(x = severity_of_harassment, fill = type_of_account)) +
  geom_bar() +
  theme_minimal() +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme(axis.text.x = element_text(angle=45, hjust = 1, size = 9)) + 
  theme(axis.text.y = element_text(size = 6)) +
  labs(x = "Severity of Harassment", y = "Number of Tweets", fill = "Type of Account") +
  theme(legend.position = "bottom") +
  theme(legend.text = element_text(size = 6)) +
  theme(legend.title = element_text(size = 9)) +
  scale_fill_taylor(palette = "lover",
  discrete = TRUE,
  reverse = TRUE)

```

## Catherine McKenna
MP McKenna chose not to seek re-election during the 2021 Federal Election and has since returned to public life. Although she cites wanting to help other countries around the world fight climate change as her main reason for leaving, the online and offline harassment she has and continues to face surely impacted her decision. Within the small sample of Tweets used in this trial study, MP McKenna's were found to be the most harassing, with 270 (54%) questioning her authority and 153 (30.6%) employing name-calling/gender insults (see @tbl-mckenna_severity). 13 Tweets (2.6%) were found to either contain vicious language, credible threats, or hate speech, which is a larger number then the 10 Tweets (2.0%) that were positive. 54 (10.8%) of the Tweets were determined to be neutral - a much smaller percent then MP May and MP Collins. 

## Elizabeth May
Unlike MP McKenna, the Tweets sent to MP May were more neutral in tone and were interesting in debating policies surrounding climate change.  

## Laurel Collins 
Of the 13 Tweets in MP Collins' sample, 11 (84.6%) were neutral in tone, discussing NDP platform commitments and events MP Collins attended on the campaign trail. The other 2 Tweets (15.4%) unsurprisingly questioned her authority, a likely harbinger of the severity of harassment that will be detected in the full-scale version of this study.  

“Few studies have specifically considered the role of status—the visibility or public profile of a politician—as a condition for the occurrence of negativity toward women”(Rheault et al., 2019, p. 2)

## Deliberative Democracy
Referencing previous discussions of the characteristics of what constitutes a healthy deliberative democracy, findings from this trial study will be compared and contrasted to evaluate the extent at which the discourse is healthy for Canadian democracy. 

The sample of MP May's Tweets analysed for this study most closed matched the characteristics of healthy discourse for democracy. 

The harassment of female politicians who communicate about climate change, one of the most crucial areas of policymaking, challenges the core values of Canada’s deliberative democracy by trying to minimize the diverse range of opinions and politicizes the subject matter, therefore risking the future of public discourse online and offline. 

## Limitations
The context in which these data sets were initially collected, focusing on mentions the three MPs received, means that my analysis and ability to draw conclusions is limited (further discussed in [Appendix A]). Due to the focus on mentions, I chose not to remove mentions during the first cleaning stage where the three MPs mentioned themselves, especially because it meant there would be even fewer Tweets from the MP Collins data set to analyze.

## Future Research
Given the changes to Twitter's affordances and the increasingly polarized political climate in Canada following the Covid-19 pandemic, there are multiple aspects of this trial study that will differ in the full-scale version of this study. MP Collins has raised her profile, through her national work as the NDP's Environment Critic and local work as the MP for Victoria, so it is anticipated that there will be more Tweets of hers to analyse and the severity of harassment will fall higher on the seven-point scale.

Some of my approaches, including how type of account is determined need to be refined ahead of the full scale version of this project. 

# Conclusion
In this trial study, I examined Tweets explicitly discussing climate change and mentioning former MP Catherine McKenna, MP Elizabeth May, and MP Laurel Collins to see the severity of harassment on a seven-point scale and type of accounts behind the harassment. My study reveals that over 75% 

# Appendix A

This datasheet provides background on who originally collected the raw Twitter data used in this trial study, how it has been previously used, and aims to reveal some of the limitations imposed on this study. The questions below are amended from the ones outlined in @gebru2021datasheets. 

## Motivation
1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*

    - This dataset was created for a group project under the theme “who follows and mentions who during elections” for        CMN 4160 - Digital Social Research taught by Professor Elizabeth Dubois at the University of Ottawa. 

    - Inessa De Angelis was the group member who led research project design and pitched the idea of selecting former MP Catherine McKenna, MP Elizabeth May, and MP Laurel Collins because of their engagement in climate change policy advocacy. 

    - Because of her previous work with Catherine McKenna during the 2019 Federal Election Campaign, Inessa noticed the connection between climate change denial, misogyny, and online harassment. After reviewing relevant literature from both Canada and other countries, Inessa noticed a large gap and wanted to help fill the gap through this project and her continued research at the Faculty of Information at the University of Toronto. 

    - The dataset specifically collected mentions of the three female politicians, not simply replies, which differs from how Inessa will approach the full-scale version of this project. Instead of quantitative content analysis, the group performed social network analysis using Gephi on the datasets. 

2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*

    - Inessa defined the parameters on behalf of her group of what data they wanted collected and Trevor Deley, PhD Candidate in E-Business (supervised by Professor Dubois) collected the data. Both were students at the University of Ottawa. 

3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*

    - Trevor's financial situation during the creation of this dataset is unknown.

    - Inessa’s tuition was partially covered by a University of Ottawa Dean’s Honour List Merit Scholarship, but there was no direct grant associated with this project.

4. *Any additional comments?*

    - While Inessa has previously worked with former MP Catherine McKenna during the 2019 Federal Election campaign, it is important to note that she did not have access to MP McKenna's social media accounts nor produce social media content, including the Tweets in the data set. 
    
    - Since the creation of this dataset in 2021 and the inital version of this project was completed, MP McKenna has left public life and pursued other projects. This has no impact on the composition of this data set, but is important context. 

## Composition

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)?* 
    
    - Each row of the dataset is an individual. However, not every row is a unique individual because some people sent multiple Tweets to the MPs.   
    
2. *How many instances are there in total ?*  

    - Between Catherine McKenna, Elizabeth May, and Laurel Collins there were 116,649 Tweets in the raw datasets.   

3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not*  

    - No, the raw data set does not contain all the mentions of MPs McKenna, May, and Collins on Twitter leading up to and during the 2019 Federal election campaign but it is also not a sample. Tweets in the data set range from early August to late October, with the election campaign running from September 11 - October 21, 2019. The missing instances is due to accounts being suspended or deleted from the platform and accounts being set to "private" before data collection happened in early 2021. 
    
    - Given the affordances of Twitter, the sample of Tweets is representative of the nature of discourse about climate policy on Twitter before/during the election campaign because there is diversity in the age, gender, politicial affiliation, and geographic coverage (verified through user bios, locations, and other personal information attatched to their public profiles). 
    
4. *What data does each instance consist of? “Raw” data (for example, unprocessed text or images) or features? In either case, please provide a description.*

    - Each instance contains raw Tweet text and personal account information (that is displayed on their public profile) such as location, bio, username, and display name. 
    
5. *Is there a label or target associated with each instance?*

    - No, there is not.         

6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable).*

    - Yes. Many individuals opted not to list a location on their profile or write a bio, which makes it unavailable and listed as NA in the raw data set. Not sharing a location or writing a bio could be due to individuals wanting to protect their privacy or not reveal any personal information about themselves. 

7. *Are relationships between individual instances made explicit? If so, please describe how these relationships are made explicit.*

    - Yes, each individual instance mentions at least one of the three female politicians of study: former MP McKenna, MP May, or MP Collins and the word "climate" is mentioned somewhere in the Tweet.

8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*

    - No.

9. *Are there any errors, sources of noise, or redundancies in the dataset?*
 
     -  It has been over four years since these Tweets were initally sent and over two years since these Tweets were         collected for the dataset, so there is potential inaccuracies in reporting which accounts are still on the platform and which have been suspended/deleted.
     - Some users do not have bios or locations indicated (listed as NA), which may impact how coders categorized "type of account".
    
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources?*
    
    - It is self-contained. 
    
11. *Does the dataset contain data that might be considered confidential* 

    - No, all data was gathered from publicly available sources. 

12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.* 

    - Yes, some people may find the nature of the Tweets directed toward the female politicians offensive and harmful, between the vicious language, threats, and climate mis- and dis-information. Viewers are encouraged to take care and seek assistance if required. 

13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset*
  
    - No, the data set does not have any columns that directly identify any sub-populations, but information regarding gender, location, politicial ideology and party affiliation can be gleaned from the "location" and "bio" columns. 
    
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset?*

    - Yes, some individuals can be identified by their username, display name, profile picture, and bio while others have opted to maintain anonymous profiles. 

15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*

    - Yes, the data set contains sensitive information, such as political party affiliation, ideology, and location, however this is public information as users list it in their bio or on the House of Common's website in the instance of the three female politicians.

16. *Any other comments?*

    - No further comments.

## Collection Process

1. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*

    -  Trevor collected these Tweets using a custom Python script to access Twitter’s API. The Tweets were also run         through Perspective API’s sentiment detection software to check for toxicity and assign a score, although Inessa removed these scores during the initial cleaning process for this paper. 
    
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*

    - Tweets were collected via Twitter's API using Python. 
    
    - Inessa is unsure if Trevor employed any validation mechanisms or procedures. 

3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*

    - The data set is not a sample, it contains all the Tweets that were publicly available when Trevor collected the data.
    
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*

    - Trevor Deley (compensation unknown).
    
    - Professor Elizabeth Dubois (Associate Professor salary from the University of Ottawa, but not directly for this       project).
    
    - Inessa De Angelis (no compensation, paid tuition to the University of Ottawa to take CMN 4160).

5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*

      - Tweets were collected once, in late January/early February 2021 (a specific day/time is unknown). 
      
      - The Tweets included in the data set were created between the start of August and the end of October of 2019. 

6. *Were any ethical review processes conducted (for example, by an institutional review board)?*

    - No.

7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*

    - All data was obtained via a third party: Twitter.

8. *Were the individuals in question notified about the data collection?*

    - No.
    
9. *Did the individuals in question consent to the collection and use of their data?*

    - No.

10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses?*

    - No, consent was never obtained.

11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted?*

    - No, an analysis of this nature has not been conducted. 

12. *Any other comments?*

    - No further comments.

## Cleaning/Labelling 

1. *Was any preprocessing/cleaning/labeling of the data done?*

    - It is unknown whether Trevor conducted any cleaning of the raw datasets with Python, beyond obtaining toxicity        scores via Perspective API. 
    
    - For this trial study, Inessa used R to clean and simplify column names, remove additional variables, and filter       out Tweets which did not contain the word `climate`.
    
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)?*

    - Yes. Inessa has the original raw data sets saved in her personal Google Drive and on the desktop of her personal       Mac. 
    
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*

    - Trevor used Python and Perspective API by Jigsaw. Perspective API is currently free to use but requires                registration.
    
    - Inessa uses R. 

4. *Any other comments?*

    - No further comments.
    
## Uses

1. *Has the dataset been used for any tasks already? If so, please provide a description*

    - Yes. This dataset was used for the aforementioned CMN 4160 class project. 
    
    - We used Gephi to conduct social network analysis to see which communities formed around the Tweets and Twitter        accounts of former MP Catherine McKenna, MP Elizabeth May, and MP Laurel Collins. We found that McKenna had 26        communities, May had 131 communities, and Collins had 18. We also found that harssment was sent by users with         personal accounts and was most likely to show up in the same community as the three female politicans. These          findings were written up in a professional-style guide, aimed at women running for office. This project report        was used by Professor Dubois as an example for future students in her class, but not publicly distributed beyond      the class’ Brightspace page. 

2. *Is there a repository that links to any or all papers or systems that use the dataset?*
  
    - No.

3. *What (other) tasks could the dataset be used for?*
  
    - Beyond this trial study, this data set could be used to see the amount of discourse about other policy issues that mattered before/during the 2019 Federal election, such as the economy and housing.

4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*

    - No.

5. *Are there tasks for which the data set should not be used? If so, please provide a description.*

    - This data set (and any Twitter data generally) should not be used to measure public opinion, as it is not representative of the general Canadian public. 

6. *Any other comments?*

    - No further comments.

## Distribution

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created?* 

    - No. Due to Twitter’s policies about sharing data, this dataset is not publicly available but can be obtained by contacting Inessa De Angelis (inessa.deangelis@mail.utoronto.ca). Twitter’s policies outline that datasets are not to be uploaded to GitHub or similar sites/repositories without removing all identifying information and only including the Tweet ID or User ID. 

2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*

    - No, this data set will not be distributed and does not have a DOI. 

3. *When will the dataset be distributed?*
    
    - The dataset is currently available. Please contact Inessa via email to access it. 

4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)?* 

    - No.
    
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances?*

    - No restrictions are known.
    
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances?*

    - No restrictions are known.

7. *Any other comments?*

    - No further comments.

## Maintenance

1. *Who will be supporting/hosting/maintaining the dataset? How can the owner/curator/manager of the dataset be contacted?*

    - There is no regular maintaince of this data set, but Inessa De Angelis is responsible. 
    
    - She can be reached via email at inessa.deangelis@mail.utoronto.ca. 
    
# Appendix B
The coding form is available here: https://forms.gle/ryyv4YFKKmpfTs4c6 

## Form Preamble 
This is the coding form for the Torrential Twitter Trial Study. Please refer to the Codebook (available in the GitHub repository for this project under Inputs --> Codebook) while coding Tweets. 

Please contact inessa.deangelis@mail.utoronto.ca if you have questions or require any further information.

## Form Questions
(@) Coder name

    - Inessa

(@) Name of Politician

    - Catherine McKenna
    
    - Elizabeth May
    
    - Laurel Collins 

(@) Tweet URL

    - (short text answer)

(@) Tweet text

    - (long text answer)
  
(@) Username

    - (short text answer)
    
(@) Name

    - (short text answer)  
    
(@) Bio

    - (long text answer)     
    
(@) Severity of harassment 

    - 1    
    
    - 2    
 
    - 3

    - 4

    - 5
    
    - 6
    
    - 7
    
(@) Type of account

    - 1    
    
    - 2    
 
    - 3

    - 4

    - 5
    
    - 6
    
# Appendix C

# References